---
Admissions Modeling:
Modeling how likely a person will be admitted to Olin College of Engineering based on how similar they are to already admitted Olin students based on SAT Scores/ACT Scores, high school GPA, and gender (add more?take away some?)
---

Loading in csv data on SAT Scores, ACT Scores, high school GPA, gender (only includes men or women as options), and enrollment. 
```{r}
# SAT Scores
SAT25thReading <- read.csv("25thPercentileReading.csv", header = FALSE)
SAT75thReading <- read.csv("75thPercentileReading.csv", header = FALSE)
SAT25thMath <- read.csv("25thPercentileMath.csv", header = FALSE)
SAT75thMath <- read.csv("75thPercentileMath.csv", header = FALSE)

# ACT Scores
ACT25thReading <- read.csv("ACT25thPercentileReading.csv", header = FALSE)
ACT75thReading <- read.csv("ACT75thPercentileReading.csv", header = FALSE)
ACT25thMath <- read.csv("ACT25thPercentileMath.csv", header = FALSE)
ACT75thMath <- read.csv("ACT75thPercentileMath.csv", header = FALSE)

# Percent of Oliners in a GPA range
GPA <-read.csv("GPA.csv", header = FALSE)

#Gender
TotalApplicants <- read.csv("TotalApplicants.csv", header = FALSE)
ApplicantsMen <- read.csv("ApplicantsMen.csv", header = FALSE)
AdmittedMen <- read.csv("AdmittedMen.csv", header = FALSE)
ApplicantsWomen <- read.csv("ApplicantsWomen.csv", header = FALSE)
AdmittedWomen <- read.csv("AdmittedWomen.csv", header = FALSE)

# Enrollment
Enrollment <- read.csv("Enrollment.csv", header = FALSE)

```
Making data usable to analyze by organizing them, converting them to numeric if needed, and cleaning out NA's. 
```{r}
# converting dataframes to numerics and removing NA's
LowerSATReading <- as.numeric(SAT25thReading[2,3:length(SAT25thReading)])
LSR<-LowerSATReading[!is.na(LowerSATReading)]

HigherSATReading <-as.numeric(SAT75thReading[2,3:length(SAT75thReading)])
HSR <- HigherSATReading[!is.na(HigherSATReading)]

LowerSATMath<- as.numeric(SAT25thMath[2,3:length(SAT25thMath)])
LSM <- LowerSATMath[!is.na(LowerSATMath)]

HigherSATMath <-as.numeric(SAT75thMath[2,3:length(SAT75thMath)])
HSM <- HigherSATMath[!is.na(HigherSATMath)]

LowerACTReading <- as.numeric(ACT25thReading[2,3:length(ACT25thReading)])
LAR <- LowerACTReading[!is.na(LowerACTReading)]

HigherACTReading <- as.numeric(ACT75thReading[2,3:length(ACT75thReading)])
HAR <- HigherACTReading[!is.na(HigherACTReading)]

LowerACTMath <- as.numeric(ACT25thMath[2,3:length(ACT25thMath)])
LAM <- LowerACTMath[!is.na(LowerACTMath)]

HigherACTMath <- as.numeric(ACT75thMath[2,3:length(ACT75thMath)])
HAM <- HigherACTMath[!is.na(HigherACTMath)]

TotalApplicant <- as.numeric(TotalApplicants[2,3:length(TotalApplicants)])
TotalApp <- TotalApplicant[!is.na(TotalApplicant)]

ApplicantMen <- as.numeric(ApplicantsMen[2,3:length(ApplicantsMen)])
AppMen <- ApplicantMen[!is.na(ApplicantMen)]

ApplicantWomen <- as.numeric(ApplicantsWomen[2,3:length(ApplicantsWomen)])
AppWomen <- ApplicantWomen[!is.na(ApplicantWomen)]

AdmitMen <- as.numeric(AdmittedMen[2,3:length(AdmittedMen)])
AdmMen <- AdmitMen[!is.na(ApplicantMen)]

AdmitWomen <- as.numeric(AdmittedWomen[2,3:length(AdmittedWomen)])
AdmWomen <- AdmitWomen[!is.na(AdmitWomen)]

PercentHighGPA <- as.numeric(GPA[2, 2:length(GPA)])
PercentMediumGPA <- as.numeric(GPA[3, 2:length(GPA)])
PercentLowGPA <- as.numeric(GPA[4, 2:length(GPA)])

Enrolled <- as.numeric(Enrollment[2,3:length(Enrollment)])
Enroll <- Enrolled[!is.na(Enrolled)]
```


Means of Gender Data
```{r}
# average number of applicants and admitted students each year
avgTotalApp <- mean(TotalApp)
avgAppMen <- mean(AppMen)
avgAppWomen <- mean(AppWomen)
avgAdmMen <- mean(AdmMen)
avgAdmWomen <-mean(AdmWomen)

# probability admitted based on gender
ProbAdmMen <- avgAdmMen/avgAppMen
ProbAdmWomen <- avgAdmWomen/avgAppWomen
ProbAdmGender <- c(ProbAdmMen, ProbAdmWomen)

# average of total admitted students 
avgAdm <- avgAdmMen + avgAdmWomen

# overall probability of being admitted
POverallAdm <- avgAdm/avgTotalApp
```

Enrollment Data
```{r}
# average number of students enrolled at Olin
avgEnrolled <- mean(Enroll)
avgNotEnrolled <- avgAdm - avgEnrolled

# probability of being admitted and decided to go and not to go to Olin
PAdmEnrolled <- avgEnrolled / avgAdm
PAdmNotEnrolled <- avgNotEnrolled / avgAdm
```

Mean Lows and Highs of SAT Scores
```{r}
# finding the mean high and low scores for the reading and math SAT
avgLSR <- mean(LSR)
avgHSR <- mean(HSR)
avgLSM <- mean(LSM)
avgHSM <- mean(HSM)

# average low and high SAT composite scores of Oliners
avgLSC <- avgLSR + avgLSM
avgHSC <- avgHSR + avgHSM

SATScores <- c(avgLSR, avgHSR, avgLSM, avgHSM)
SATComp <- c(avgLSC, avgHSC)

print(SATScores)
```
Using Bayes' Rule to Find Probability of Being Admitted Given Within a GPA Range

```{r}
# probabilities Oliners are SAT score ranges

# 75th percentile: 1563 or greater
PHighSATOlin <- 0.25

# 25th percentile: 1433 to 1563
PMediumSATOlin <- 0.5

# 0th percentile: 1300 to 1433
PLowSATOlin <- 0.25


# probability high schooler being within these SAT ranges from national data
PHighSAT <- 0.03
PMediumSAT <- 0.05
PLowSAT <- 0.07

# function to determine probability of being admitted given your SAT score
PSATAdm <- PMediumSATOlin*PAdmEnrolled + PMediumSATOlin*PAdmNotEnrolled
PAdmSAT <- function(YourSAT)
  if (YourSAT>avgHSC){
    PAdmSAT <- PHighSATOlin*POverallAdm/PHighSAT
  } else if(YourSAT>avgLSC){
    PAdmSAT <- PMediumSATOlin*POverallAdm/PMediumSAT
  } else if(YourSAT>1300){
    PAdmSAT <- PLowSATOlin*POverallAdm/PLowSAT
  } else {
    PAdmSAT <-0
  }


PSATAdm*POverallAdm/PMediumSAT
```

Mean Lows and Highs of ACT Scores
```{r}
# mean highs and lows of Reading and Math ACT scores
avgLAR <- mean(LAR)
avgHAR <- mean(HAR)
avgLAM <- mean(LAM)
avgHAM <- mean(HAM)

# average high and low ACT composite scores of Oliners
avgOlinLAC <- (avgLAR + avgLAM) / 2
avgOlinHAC <- (avgHAR + avgHAM) / 2


ACTScores <- c(avgLAR, avgHAR, avgLAM, avgHAM)
ACTComp <- c(avgOlinLAC, avgOlinHAC)
print(ACTComp)

```

Using Bayes' Rule to Find Probability of Being Admitted into Olin based on being in an ACT Range
```{r}
# probabilities Oliners are in ACT score ranges

# 75th percentile: 34 or greater
PHighACTOlin <- 0.25

# 25th percentile: 32.7 to 34
PMediumACTOlin <- 0.5

# 0th percentile: 30 to 32.7
PLowACTOlin <- 0.25

# probability of a higher schooler being within these ACT score ranges
PHighACT <- 0.01
PMediumACT <- 0.04
PLowACT <- 0.05

PAdmACT <- function(YourACT)
  if (YourACT>avgOlinHAC){
    PAdmACT <- PHighACTOlin*POverallAdm/PHighACT
  } else if(YourACT>avgOlinLAC){
    PAdmACT <- PMediumACTOlin*POverallAdm/PMediumACT
  } else if(YourACT>30){
    PAdmACT <- PLowACTOlin*POverallAdm/PLowACT
  } else {
    PAdmACT <-0
  }
print(PAdmACT(34))
```

Means of GPA Data
```{r}
# average percent of Olin students that fall with in a GPA range
avgOlinHighGPA <- mean(PercentHighGPA)
avgOlinMediumGPA <- mean(PercentMediumGPA)
avgOlinLowGPA <- mean(PercentLowGPA)

# found from national data
avgHighGPA <- 0.07
avgMediumGPA <- 0.14
avgLowGPA <- 0.31

# probability of having high (3.74 - 4.0), medium (3.5-3.74) and low (3.25-3.49) GPAs given the student is at Olin
avgOlinGPA <- c(avgOlinHighGPA, avgOlinMediumGPA, avgOlinLowGPA)

```

Using Bayes' Rule to Find Probability of Being Admitted Given Within a GPA Range
```{r}
PAdmGPA <- function(YourGPA)
  if (YourGPA>3.74){
    PAdmGPA <- avgOlinHighGPA*POverallAdm/avgHighGPA
  } else if(YourGPA>3.49){
    PAdmGPA <- avgOlinMediumGPA*POverallAdm/avgMediumGPA
  } else if (YourGPA>3.24) {
    PAdmGPA <- avgOlinLowGPA*POverallAdm/avgLowGPA
  } else {
    PAdmGPA <-0
  }
print(PAdmGPA(3.9))
```