---
title: "Predicting Franklin W. Olin College of Engineering's Yield Rate over time"
output: html_notebook
---

## R Notebook Basics - Delete Later

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

## Introduction

In the competitive world of college admissions, many prospective students wanting to apply to the higher echelon of universities often think about many factors such as their standardized test scores, their GPAs, their recommendation letters and their essays. And while all these factors are important for a student's chances of entering their dream university, most of the times, it is also the goal of the university to make sure they accept the right students to build a well-balanced class.

A college's yield rate is the percentage of students after having been offered admission calculated by dividing the number of students who choose to enroll at a school in a given year, which is often based on their decision to pay a deposit, by the total number of offers of acceptance sent.

While a higher yield rate does indicate a sign of a more selective college, the statistic is more important for a college to decide how many students to accept in their entering freshman class. If a college miscalculates the percentage of students who accept an offer of admission, they run the risk of either underclassing or overclassing their incoming freshman class. If the university underclasses, they lose the opportunity on tuition money the university could have gained and used to improve their facilities. A university overclassing, on the other hand, may lead to losing out on potential tuition money and a poorer quality of life for their students: an issue that universities like [Northeastern](https://huntnewsnu.com/68925/editorial/op-ed-northeastern-needs-to-fix-the-current-housing-crisis/) and [schools in the University of California](https://www.latimes.com/california/story/2022-09-26/college-housing-shortage-pushes-students-into-crisis-as-most-uc-classes-start-up) system faces now.

And with those risks, every college, since it is essential to keep students happy while maximizing profit, should examine themselves, in our analysis of the Olin College of Engineering, and level up their expectations for their yield rates before going into the next recruitment season. 

## Setup

### Load Data

Loading in admissions data (pre-formatted in Excel)

Variables Include
- Percent Admitted (Total, Men, Women)
- Percent Admitted Yield (Total, Men, Women)
- Number of Applicants (Total, Men, Women)
- Number of Admitted (Total, Men, Women)
- Number Enrolled (Total, Men, Women)

### Convert Data into Dataframe

```{r}
admissions <- read.csv("yield-rate-data.csv")
admissions
```

### Format Total Dataframe into TimeSeries

```{r}
admissions_ts <- ts(admissions[,-1], start=c(2006), end=c(2021),frequency=1)
admissions_ts
```

### Select Admissions Yield Total and verify if output is timeseries

```{r}
admissions_yield_ts <- ts(admissions_ts[,4], start=c(2006), end=c(2021),frequency=1)
admissions_yield_ts
print(admissions_yield_ts)
```

## General Analysis

### Get Summary Statistics of Admissions Timeseries

```{r}
summary(admissions_yield_ts)
```

### Analyze General Trend with a Linear Regression Line

```{r}

ts.plot(admissions_yield_ts, xlab="Year", ylab="Yield Percentage", main="Olin College Yield Percentage Over Time, 2006-2021", ylim=c(50,80), col = 'black') 
abline(reg=lm(admissions_yield_ts~time(admissions_yield_ts)), col='red')
legend("topleft",
       legend = c("Total Yield Rate", "Fitted Line"),
       col = c('black', 'red'),
       pch = 16)

```

## Auto-Correlation

To determine what type of series we have, whether it is a white noise, random walk, auto regressive or moving average.

```{r}
acf(admissions_yield_ts, main="Autocorrelation of Total Yield Percentage")
```

## ARIMA Modelling

Create AutoRegressive (AR) Model

```{r}
AR <- arima(admissions_yield_ts, order = c(1,0,0))
print(AR)
```

 
```{r}
ts.plot(admissions_yield_ts, ylab = "Yield Percentage", main="Fitted Auto Regressive Admission Yield Model Over Time",col='black',ylim=c(50,80))
AR_fit <- admissions_yield_ts - residuals(AR)
points(AR_fit, type = "l", col = 'red', lty = 2)
legend("topleft",
       legend = c("Total Yield Rate", "Fitted Line"),
       col = c('black', 'red'),
       pch = 16)
```

Using predict() to make a 1-step forecast and to make 1-step through 10-step forecasts

```{r}
predict_AR <- predict(AR)

predict(AR, n.ahead = 10)
```

Plotting the series plus the forecast and 95% prediction intervals

```{r}
ts.plot(admissions_yield_ts, ylab = "Yield Percentage", main = "AR Predicted Yield Over One Year", col = 'black', xlim = c(2006, 2022), ylim = c(50, 80))
AR_forecast <- predict(AR, n.ahead = 10)$pred
AR_forecast_se <- predict(AR, n.ahead = 10)$se
points(AR_forecast, type = "l", col = 'red')
points(AR_forecast - 2*AR_forecast_se, type = "l", col = 'blue', lty = 2)
points(AR_forecast + 2*AR_forecast_se, type = "l", col = 'blue', lty = 2)
legend("topleft",
       legend = c("Total Yield Rate", "Fitted Prediction", "95% Confidence Interval"),
       col = c('black', 'blue', 'red'),
       pch = 16)
```

Fitting the MA model to admissions_total_yield

```{r}
MA <- arima(admissions_yield_ts, order = c(0,0,1))
print(MA)
```

Plotting the series along with the MA fitted values

```{r}
ts.plot(admissions_yield_ts, ylab = "Yield Percentage", main="Fitted Moving Average Admission Yield Model Over Time",col='black',ylim=c(50,80))
MA_fit <- admissions_yield_ts - residuals(MA)
points(MA_fit, type = "l", col = 'red', lty = 2)
legend("topleft",
       legend = c("Total Yield Rate", "Fitted Line"),
       col = c('black', 'red'),
       pch = 16)
```

Making and obtaining a 1-step forecast based on MA as well as a 10-step forecast

```{r}
predict_MA <- predict(MA)
predict_MA$pred[1]

predict(MA,n.ahead=10)
```

Plotting the Total Admissions Yield Percentage data with the forecast and 95% prediction intervals

```{r}
ts.plot(admissions_yield_ts, ylab = "Yield Percentage", main = "MA Predicted Yield Model Over One Year", col = 'black', xlim = c(2006, 2022), ylim = c(50, 80))
MA_forecasts <- predict(MA, n.ahead = 10)$pred
MA_forecast_se <- predict(MA, n.ahead = 10)$se
points(MA_forecasts, type = "l", col = 'red')
points(MA_forecasts - 2*MA_forecast_se, type = "l", col = 'blue', lty = 2)
points(MA_forecasts + 2*MA_forecast_se, type = "l", col = 'blue', lty = 2)
legend("topleft",
       legend = c("Total Yield Rate", "Fitted Prediction", "95% Confidence Interval"),
       col = c('black', 'blue', 'red'),
       pch = 16)
```

## Analyzing Models

Find correlation between AR_fit and MA_fit

```{r}
cor(AR_fit, MA_fit)
```

Find AIC (Akaike Information Criterion) of AR and MA Models

```{r}
AIC(AR)
```

```{r}
AIC(MA)
```

Find BIC (Bayesian Information Criterion) of AR and MA Models

```{r}
BIC(AR)
```

```{r}
BIC(MA)
```

## Conclusion

And how should interested applicants use this calculation to their advantage?
- Be interested in Olin
- Learn more about it

Many students make the choice, a good choice of attending Olin! <- insert propaganda

## Resources

- https://en.wikipedia.org/wiki/Yield_(college_admissions)#:~:text=It%20is%20calculated%20by%20dividing,particular%20school%20of%20higher%20education.
- https://www.rdocumentation.org/packages/tseries/versions/0.7-5/topics/arma
- https://rpubs.com/JSHAH/481706
- https://otexts.com/fpp2/arima-r.html
- https://www.statology.org/aic-in-r/
- https://www.statology.org/bic-in-r/
