---
title: "Olin College's Yield Rate over Time"
output: html_notebook
---

## R Notebook Basics

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

## Introduction

Analyzing Olin College's Yield Rate over time.

Yield Rate is how many students accepted by college / how many students get accepted.

## Load Data

Loading in admissions data (pre-formatted in Excel)

Variables Include
- Percent Admitted (Total, Men, Women)
- Percent Admitted Yield (Total, Men, Women)
- Number of Applicants (Total, Men, Women)
- Number of Admitted (Total, Men, Women)
- Number Enrolled (Total, Men, Women)

```{r}
admissions <- read.csv("yield-rate-data.csv")
admissions
```

```{r}
admissions_yield <- admissions$Admissions.yield...total
admissions_yield
```

## Admissions Yield Over Time

```{r}
p <- ggplot(aes(x = Year, y = Admissions.yield...total), data = admissions) + geom_point()

p
```












